---
output: github_document
always_allow_html: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# quanteda.llm

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/quanteda.llm)](https://CRAN.R-project.org/package=quanteda.llm)
[![R-CMD-check](https://github.com/quanteda/quanteda.llm/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/quanteda/quanteda.llm/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/quanteda/quanteda.llm/graph/badge.svg)](https://app.codecov.io/gh/quanteda/quanteda.llm)
[![pkgdown](https://img.shields.io/badge/pkgdown-site-blue)](https://quanteda.github.io/quanteda.llm/)
<!-- badges: end -->

The **quanteda.llm** package makes it easy to use LLMs with [quanteda](https://quanteda.io/) corpora (or character vectors and data frames), to enable classification, summarisation, scoring, and analysis of documents and text. **quanteda** provides a host of convenient functions for managing, manipulating, and describing corpora as well as linking their document variables and metadata to these documents.  **quanteda.llm** makes it convenient to link these to LLMs for analysing or classifying these texts, creating new variables from what is created by the LLMs. Using a tidy approach and linking to the new **quanteda.tidy** package, we enable convenient operations using common Tidyverse functions for manipulating LLM-created objects and variables.



# Included functions

The package includes the following functions:

- `ai_text()`: 
  - A generic function that can be used with any LLM supported by
`ellmer`.
  - Generates structured responses or classifications based on pre-defined instructions
for texts in a `quanteda corpus`. 
  - Users can flexibly define prompts and structure of responses via 
  `type_object()` from the [`ellmer` package](https://ellmer.tidyverse.org/articles/structured-data.html).
  - Users can add a dataset with examples to improve LLM performance (few-shot prompting)
  - Supports resuming interrupted processes in a `result_env` environment.
  
- `ai_validate()`: 
  - Starts an interactive app to manually validate the
  LLM-generated outputs.
  - Allows users to review and validate the LLM-generated outputs and justifications, marking them as valid or invalid.
  - Supports resuming the validation process in case of interruptions in a `result_env` environment.
  
- `ai_summary()`:
  - A wrapper around `ai_text()` for summarizing documents in a corpus.
  - Uses a pre-defined `type_object()` to structure the summary output.
  
- `ai_salience()`:
  - A wrapper around `ai_text()` for computing salience scores for topics in a corpus.
  - Uses a pre-defined `type_object()` to structure the salience classification output.
  
- `ai_score()`:
  - A wrapper around `ai_text()` for scoring documents based on a scale defined by a prompt.
  - Uses a pre-defined `type_object()` to structure the scoring output.

# Supported LLMs

The package supports all LLMs currently available with the `ellmer`
package, including:

- Anthropic’s Claude: `chat_anthropic()`.
- AWS Bedrock: `chat_aws_bedrock()`.
- Azure OpenAI: `chat_azure_openai()`.
- Cloudflare: `chat_cloudflare()`.
- Databricks: `chat_databricks()`.
- DeepSeek: `chat_deepseek()`.
- GitHub model marketplace: `chat_github()`.
- Google Gemini/Vertex AI: `chat_google_gemini()`,
  `chat_google_vertex()`.
- Groq: `chat_groq()`.
- Hugging Face: `chat_huggingface()`.
- Mistral: `chat_mistral()`.
- Ollama: `chat_ollama()`.
- OpenAI: `chat_openai()`.
- OpenRouter: `chat_openrouter()`.
- perplexity.ai: `chat_perplexity()`.
- Snowflake Cortex: `chat_snowflake()` and `chat_cortex_analyst()`.
- VLLM: `chat_vllm()`.

For authentication and usage of each of these LLMs, please refer to the
respective `ellmer` documentation
[here](https://ellmer.tidyverse.org/reference/index.html). **For
example,** to use the `chat_openai` models, you would need to sign up for an
API key from [OpenAI](https://platform.openai.com/playground/prompts) which you can save in your `.Renviron` file as
`OPENAI_API_KEY`. To use the `chat_ollama` models, first download and install
[Ollama](https://ollama.com/). Then install some models either from the
command line (e.g. with ollama pull llama3.1) or within R using the
`rollama` package. The Ollama app must be running for the models to be
used. 

## Installation

You can install the development version of **quanteda.llm** from
<https://github.com/quanteda/quanteda.llm> with:

```{r, eval = FALSE}
# install.packages("pak")
pak::pak("quanteda/quanteda.llm")
pak::pak("quanteda/quanteda.tidy")
```

## Example use

To learn more about how to use the package, please refer to the following examples:

- [Summarizing texts with LLMs](https://quanteda.github.io/quanteda.llm/articles/pkgdown/examples/summarizing.html)
- [Salience ratings of topics in texts with LLMs](https://quanteda.github.io/quanteda.llm/articles/pkgdown/examples/salience.html)
- [Scoring texts with LLMs](https://quanteda.github.io/quanteda.llm/articles/pkgdown/examples/scoring.html)
- [Structuring LLM responses for text analysis](https://quanteda.github.io/quanteda.llm/articles/pkgdown/examples/structuring.html)
- [Validating LLM responses](https://quanteda.github.io/quanteda.llm/articles/pkgdown/examples/validating.html)



