---
title: "Example: Salience rating of topics"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The `ai_salience()` function allows you to classify documents based on their relevance to predefined topics. The function uses a predefined `type_object` argument from `ellmer` to structure the LLM's response, producing a list of topics and their salience scores for each document. This function is particularly useful for analysing large corpora where manual classification would be impractical. Users need to provide a character vector of documents and a list of topics to classify. The LLM will then analyse each document and assign a salience score to each topic, indicating how relevant the document is to that topic.

### Loading packages and data

```{r getting-started}
library(quanteda)
library(quanteda.llm)

data_corpus_inaugural <- data_corpus_inaugural[57:60]
```

### Using `ai_salience()` for salience rating of topics

```{r ai_salience}
# define the topics for salience classification
topics <- c("economy", "environment", "healthcare")
result <- data_corpus_inaugural %>%
  ai_salience(topics, chat_fn = chat_openai, model = "gpt-4o",
              api_args = list(temperature = 0, seed = 42))
```

```{r display_results1, eval = FALSE}
result
```

```{r display_results2, echo = FALSE}
# topic and score are created as new docvars in the corpus
library(kableExtra)
result %>%
  kable("html", escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  column_spec(3) 
```
