---
title: "Example: Scaling texts"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The `ai_score()` function allows you to score documents based on a predefined scale. The function uses a predefined `type_object` argument from `ellmer` to structure the LLM's response, producing a score for each document based on the specified scale as well as a short justification for the score. This function is useful for evaluating documents against specific criteria or benchmarks. Users need to provide a character vector of documents and a scale to score against. The LLM will then analyse each document and assign a score based on the provided scale, along with a brief explanation of the reasoning behind the score.

### Loading packages and data

```{r getting-started}
library(quanteda)
library(quanteda.llm)
library(quanteda.tidy)

data_corpus_inaugural <- tail(data_corpus_inaugural, 3)
```

### Using `ai_score()` for scoring documents

```{r ai_score}
prompt <- "Score the following document on a scale of how much it aligns 
with the political left. The political left is defined as groups 
which advocate for social equality, government intervention in the 
economy, and progressive policies. Use the following metrics: 

SCORING METRIC: 
3 : extremely left 
2 : very left 
1 : slightly left 
0 : not at all left"

data_corpus_inaugural <- data_corpus_inaugural %>%
  mutate(ai_score(text, prompt, chat_fn = chat_openai, model = "gpt-4o",
                  api_args = list(temperature = 0, seed = 42)))

# score and evidence are created as new docvars in the corpus
library(kableExtra)
cbind(data.frame(docnames = docnames(data_corpus_inaugural)), 
      select(docvars(data_corpus_inaugural), score, evidence)) %>%
  kable("html", escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  column_spec(3) 
```
