---
title: "Example: Summarizing texts"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The `ai_summary()` function allows you to summarize documents using an LLM. It generates a summary for each document in a character vector and stores it as a new character vector which can be added as a document variable in a `quanteda` corpus element. The function uses a predefined `type_object` argument from `ellmer` to structure the LLM's response, producing succinct summaries of each document. Users need to provide a character vector of documents to summarize and choose the LLM provider they want to use for summarization.

### Loading packages and data

```{r getting-started}
library(quanteda)
library(quanteda.llm)
library(quanteda.tidy)

data_corpus_inaugural <- tail(data_corpus_inaugural, 3)
```

### Using `ai_summary()` for summarization of documents

```{r ai_summary}
# creates a new docvar "summary"
data_corpus_inaugural <- data_corpus_inaugural %>%
  mutate(ai_summary(text, chat_fn = chat_openai, model = "gpt-4o",
                    api_args = list(temperature = 0, seed = 42)))

library(kableExtra)
cbind(data.frame(docnames = docnames(data_corpus_inaugural)), 
      summary = data_corpus_inaugural$summary) %>%
  kable("html", escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  column_spec(2)
```