---
title: "Example: Manually validating LLM responses"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The `ai_validate()` function allows users to manually check and validate the responses generated by the LLM with a user-friendly ShinyApp. Such manual checks are essential for ensuring the quality and accuracy of the LLM's output. The function can be used to review the scores and justifications generated by the LLM, and users can also highlight and save examples from the original texts that support the validated text classifications. The saved examples can be used for further qualitative analyses or to built a labelled dataset for fine-tuning open-source LLMs to receive improved performance on similar tasks.

### Loading packages and data

```{r getting-started}
library(quanteda)
library(quanteda.llm)
library(quanteda.tidy)

data_corpus_inaugural <- tail(data_corpus_inaugural, 3)
```

### Using `ai_validate()` to manually check LLM-generated outputs

```{r ai_validate, eval = FALSE}
data_corpus_inaugural <- data_corpus_inaugural %>%
  mutate(ai_validate(text, score, evidence))
# validated scores and comments as well as highlighted examples from texts are
# created as new docvars in the corpus 
```